The Truth is Revealed: Identity of the Fish Fry "Teti" Fishery in Eastern Cuba

### 2020-12-01 analysis/paper/paper.Rmd
almost done with the paper

@@

#if (!requireNamespace("BiocManager", quietly = TRUE))
  #  install.packages("BiocManager")

#BiocManager::install("msa")

@@

Tetis (19–28 mm standard length) were collected 400 m from the river mouth of Río Yumurí (\@ref(fig:mainMap)), eastern Cuba, in December, 2018. Samples were collected using hand nets (1 mm mesh size) 2 m from the riverbank. Specimens were anesthetized using MS 222 (Tricaine Methanesulphonate) and immediately preserved in 80% ethanol. Voucher specimens were deposited in the collection of the Acuario Nacional de Cuba (catalog number: ANC-xxxx).

@@

# Checking for the position of sequences that were not edited properly so they still have nucleotides different that A, G, C, T 
nucType <- alphabetFrequency(ClustalOAlignment)
unknownNuc <- nucType[, c(5:14)]  # this indexing only contains nucleotides that are not A, G, T, C, or N or gaps
unknownPosition <- which(unknownNuc > 0, arr.ind=T)

@@

getBestModel <- function(fastaRF){
  if (class(fastaRF)!= c("DNAbin")){
    stop("This function only takes files of class 'DNAbin'. Use package 'ape' to read the input file")
  }
  formatPhyDat <- as.phyDat(fastaRF)                #getting the class used by modelTest: phyDat or pml
  modelChoice <- modelTest(formatPhyDat, tree = NULL, model = c("JC", "K80", "HKY", "TrN", "SYM", "GTR"), multicore = TRUE, mc.cores = 4)
  modelChoiceSorted <- as.data.frame(modelChoice[order(modelChoice$AICc),])
  return(modelChoiceSorted)
}

@@

                                                        # BUILINDG THE TREE
# --- library: phangorn
env = attr(bestModel, "env")    # getting the AICc for each competing model sorted
fitStart = eval(get(bestModel$Model[which.min(bestModel$AICc)], env), env)    # initial fit for the model chosen

@@

# saving the tree in the folder 'analysis/figures/'
bsTree <- plotBS(midpoint(fit$tree), bs, p = 75, type="p", bs.col = "blue")     # fit$tree is where the optimal tree is stored
optimalTree <- plotBS(bsTree, type = "p", cex = 0.2, font = 1)
# export tree with bootstrap values as node labels
write.tree(optimalTree, here(file="analysis/data/derived_data/Sicydium_ML_tree.tre"))

@@

```{r loglikelihood, include = TRUE, results='hide'}
# this result wont be displayed in the final document
loglikelihood <- round(-1*(as.numeric(fit[1])), 2)

@@

# checking that the tree is rooted
rootedTree <- is.rooted(tree)   # package ape. showing the tree root: geom_rootpoint(color="pink", size=5) 
# getting the node number for each main clade. This info is important because those #s are the one needed to edit each main clade/species
set.seed(123)
nodeNumber <- ggtree(tree) + geom_text(aes(label=node), hjust=-.5, col = "blue", size = 1) # looks horrible but it gives the clade number

@@

```{r Sicydium_ML_tree, fig.cap = paste("ML phylogenetic tree  (-lnL =", loglikelihood,") depicting the relationships among all available COI sequences of the genus *Sicydium*. Sequences highlited in green represent samples from Cuba. Numbers on branches represent boostrap values higher than 70.")}

>>>>

if (!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager")

@@

@@

Tetis (19–28 mm standard length) were collected 400 m from the river mouth of Río Yumurí (Figure \@ref(fig:mainMap)), eastern Cuba, in December, 2018. Samples were collected using hand nets (1 mm mesh size) 2 m from the riverbank. Specimens were anesthetized using MS 222 (Tricaine Methanesulphonate) and immediately preserved in 80% ethanol. Voucher specimens were deposited in the collection of the Acuario Nacional de Cuba (catalog number: ANC-xxxx).

@@

# Checking for the position of sequences that were not edited properly so they still have nucleotides different that A, G, C, T 
nucType <- alphabetFrequency(ClustalOAlignment)
unknownNuc <- nucType[, c(5:14)]  # this subset only contains nucleotides that are not A, G, T, C, or N or gaps
unknownPosition <- which(unknownNuc > 0, arr.ind=T)

@@

getBestModel <- function(fastaRF){
  if (class(fastaRF)!= c("DNAbin")){
    stop("This function only takes files of class 'DNAbin'. Use package 'ape' to read the input file")
  }
  formatPhyDat <- as.phyDat(fastaRF)                #getting the class used by modelTest: phyDat or pml
  set.seed(123)
  modelChoice <- modelTest(formatPhyDat, tree = NULL, model = c("JC", "K80", "HKY", "TrN", "SYM", "GTR"), multicore = TRUE, mc.cores = 4)
  modelChoiceSorted <- as.data.frame(modelChoice[order(modelChoice$AICc),])
  return(modelChoiceSorted)
}

@@

                                                        # BUILINDG THE TREE
# --- library: phangorn
set.seed(123)
env = attr(bestModel, "env")    # getting the AICc for each competing model sorted
set.seed(123)
fitStart = eval(get(bestModel$Model[which.min(bestModel$AICc)], env), env)    # initial fit for the model chosen

@@

# saving the tree in the folder 'analysis/figures/'
set.seed(123)
bsTree <- plotBS(midpoint(fit$tree), bs, p = 75, type="p", bs.col = "blue")     # fit$tree is where the optimal tree is stored

set.seed(123)
optimalTree <- plotBS(bsTree, type = "p", cex = 0.2, font = 1)
# export tree with bootstrap values as node labels
write.tree(optimalTree, here(file="analysis/data/derived_data/Sicydium_ML_tree.tre"))

@@

```{r loglikelihood, include = TRUE, results='hide'}
# this result wont be displayed in the final document
loglikelihood <- round(-1*(as.numeric(fit[1])), 2)
loglikelihood

@@

# checking that the tree is rooted
rootedTree <- is.rooted(tree)   # package ape. showing the tree root: geom_rootpoint(color="pink", size=5) 
# getting the node number for each main clade. This info is important because those #s are the one needed to edit each main clade/species
nodeNumber <- ggtree(tree) + geom_text(aes(label=node), hjust=-.5, col = "blue", size = 1) # looks horrible but it gives the clade number

@@

```{r Sicydium_ML_tree, fig.cap=paste("ML phylogenetic tree (-lnL =", loglikelihood,") depicting the relationships among all available COI sequences of the genus *Sicydium*. Sequences highlited in green represent samples from Cuba. Numbers on branches represent boostrap values higher than 70.")}
